Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-05 paper/main.tex
Moved code to external files to keep main tex clean

\begin{lstlisting}[label=lst:cython_header, caption=Cython header\, echo\_hash\_h.pxd]
cdef extern from "sha3nist.h":
    ctypedef unsigned char BitSequence
    ctypedef unsigned long long DataLength

    cdef enum HashReturn:
        SUCCESS = 0
        FAIL = 1
        BAD_HASHLEN = 2

    ctypedef struct hashState:
        pass

    HashReturn Hash(int, BitSequence *data, 
                    DataLength, 
                    BitSequence *hashval)
    HashReturn Init(hashState *state, 
                    int hashbitlen)
    HashReturn Update(hashState *state, 
                      BitSequence *data, 
                      DataLength databitlen)
    HashReturn Final(hashState *state, 
                    BitSequence *hashval)
\end{lstlisting}

We then create a wrapper-class that mimics hashlib, this is shown in Listing
\ref{lst:cython_impl}. Functions declared with def er python functions, cdef are
Cython only while cpdef are functions that can be accessed from both. While
this is just the implementation for the echo candidate, the almost exact same
pattern is used for all of the other candidates.

\begin{lstlisting}[label=lst:cython_impl, caption=Cython class\, echo\_hash.pyx]
cimport echo_hash_h
cimport cython

from libc.stdlib cimport *

cdef class echo:
    '''
    A class that tries to mimic the behaviour of hashlib, 
    i.e.  keeping state so that one can update the hashing 
    procedure instead of doing it from scratch.
    '''
    cdef echo_hash_h.hashState previous_state
    cdef echo_hash_h.hashState state
    cdef int finished
    cdef int hashbitlen
    cdef list hashval

    def __init__(self, int in_hashbitlen, 
                 bytes initial=None):
        self.finished = 0
        self.hashbitlen = in_hashbitlen
        echo_hash_h.Init(&self.state, self.hashbitlen)

        if initial:
            self.update(initial)

    cpdef update(self, bytes in_data):
        cdef char* data = <char *> in_data
        cdef int data_len = len(in_data)*8

        if self.finished:
            self.finished = 0
            self.state = self.previous_state

        echo_hash_h.Update(&self.state, 
                        <echo_hash_h.BitSequence *> data, 
                        data_len)

    cpdef final(self):
        cdef echo_hash_h.BitSequence *hashval = \
        <echo_hash_h.BitSequence *> \
        malloc(self.hashbitlen*8)

        # We copy the state so that we can continue to 
        # update. This equals hashlibs functionality, 
        # but not pycryptopp.
        self.previous_state = self.state

        echo_hash_h.Final(&self.state, hashval)
        self.finished = 1

        self.hashval = [hashval[i] for i from 0 <= i \
                        < self.hashbitlen / 8]
        free(hashval)

    cpdef copy(self):
        s = echo(self.hashbitlen)
        s.state = self.state
        s.previous_state = self.previous_state
        s.finished = self.finished
        if s.finished:
            s.hashval = self.hashval

        return s

    def digest(self):
        if not self.finished:
            self.final()

        return ''.join(map(chr, self.hashval))

    def hexdigest(self):
        if not self.finished:
            self.final()

        return ''.join(['%02x' % i for i in self.hashval])
\end{lstlisting}

\subsubsection{\ac{Distutils}} We use the
\ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to build and
install our code. The alternative to doing this would be to manually transform
cython files to C-files, build the shared library from this and other relevant
C-files and then install by copying into the correct system directory where
python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file setup.py.
Relevant configuration for compiling the echo extension, and installing SHA3lib
can bee seen in Listing \ref{lst:setup_py}.

\begin{lstlisting}[label=lst:setup_py, caption=Compiling echo from \ac{Distutils}]
from distutils.core import setup, Command
from distutils.extension import Extension
from Cython.Distutils import build_ext

srcpath = 'sha3lib/'
hashpath = srcpath + 'hash_functions/'

ext_modules = [
    Extension("sha3lib.hash_functions.echo.echo_hash",
        [hashpath+"echo/echo_hash.pyx", 
         hashpath+'echo/sha3nist.c', 
         hashpath+"echo/echo.c"],
    extra_compile_args=[
        '-funroll-loops',
        '-m32',
        '-march=prescott',
        '-O2',
        '-fomit-frame-pointer']
    ),
]

setup(
    name='SHA3lib',
    version='0.1',
    description='Python bindings for SHA3-256 using Cython',
    cmdclass = {'build_ext': build_ext},
    ext_modules = ext_modules,
    packages =
        ['sha3lib',
        'sha3lib.hash_functions',
        'sha3lib.hash_functions.echo',],
)

\end{lstlisting}

\subsection{Unit tests}
% Include the use of Known Answer Tests

@@

\begin{lstlisting}[label=lst:hashutil, caption=Extract from hashutil.py of Tahoe-LAFS source.]
from pycryptopp.hash.sha256 import SHA256
...
class _SHA256d_Hasher:
    ...
    def __init__(self, truncate_to=None):
        self.h = SHA256()
        self.truncate_to = truncate_to
        self._digest = None

    def update(self, data):
        assert isinstance(data, str) # no unicode
        self.h.update(data)

    def digest(self):
        if self._digest is None:
            h1 = self.h.digest()
            del self.h
            h2 = SHA256(h1).digest()
            if self.truncate_to:
                h1 = h1[:self.truncate_to]
                h2 = h2[:self.truncate_to]
            self._digest = h2

        return self._digest
\end{lstlisting}

@@

\begin{lstlisting}[label=lst:hashutilmod, caption=Parts of hashutil.py of Tahoe-LAFS source after modification.]
from sha3lib import bmw256 as SHA256
...
class _SHA256d_Hasher:
    ...
    def __init__(self, truncate_to=None):
        self.h = SHA256()
        self.truncate_to = truncate_to
        self._digest = None

    def update(self, data):
        assert isinstance(data, str) # no unicode
        self.h.update(data)

    def digest(self):
        if self._digest is None:
            h1 = self.h.digest()
            del self.h
            if self.truncate_to:
                h1 = h1[:self.truncate_to]
            self._digest = h1

        return self._digest
\end{lstlisting}

@@

For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with the included
SHA-256.  An implication of this is that uploading of mutable files renders
impossible, since the internal verification fails. For this reason, \emph{only
immutable files} are tested and measured in this paper.

@@

%**************************************%
\chapter{Discussion}
%**************************************%

>>>>

\lstinputlisting[label=lst:cython_header, caption=Cython header\, echo\_hash\_h.pxd]
{listings/echo_hash_h.pxd}

We then create a wrapper-class that mimics hashlib, this is shown in Listing
\ref{lst:cython_impl}. Functions declared with def er python functions,
{\verb cdef } are Cython only while {\verb cpdef } are functions that can be
accessed from both. While this is just the implementation for the echo
candidate, the almost exact same pattern is used for all of the other
candidates.

\lstinputlisting[label=lst:cython_impl, caption=Cython class\, echo\_hash.pyx]
{listings/echo_hash.pyx}

\subsubsection{\ac{Distutils}}
We use the \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
build and install our code. The alternative to doing this would be to manually
transform Cython files to C-files, build the shared library from this and other
relevant C-files and then install by copying into the correct system directory
where python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file
{\verb setup.py}. Relevant configuration for compiling the echo extension, and
installing SHA3lib can bee seen in Listing \ref{lst:setup_py}.

\lstinputlisting[label=lst:setup_py, caption=Compiling echo from \ac{Distutils}]
{listings/setup.py}

\subsection{Unit tests}
% TODO: Include the use of Known Answer Tests

@@

\lstinputlisting[label=lst:hashutil, caption=Extract from hashutil.py of Tahoe-LAFS source.]
{listings/hashutil.py}

@@

\lstinputlisting[label=lst:hashutilmod, caption=Parts of hashutil.py of Tahoe-LAFS source after modification.]
{listings/hashutil_mod.py}

@@

For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with the included
SHA-256.  An implication of this is that uploading of mutable files renders
impossible, since the internal verification fails. For this reason, \emph{only
immutable files} are tested and measured in this paper. Implications of this is
discussed further in Chapter \ref{ch:discussion}.

@@

%**************************************%
\chapter{Discussion}
%**************************************%
\label{ch:discussion}
